---
title: 'Project I: Data Preparation1'
author: "Chitra karki"
date: "9/7/2021"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Instructions:** While discussions with classmates are allowed and encouraged, please try to work on the
project independently and direct your questions to the TA/instructor. Please interpret your analysis results
using concise and clear language and focusing on interesting findings. You are requested to use R
Markdown to solve this project. Submissions are to be uploaded as a single (knitted) PDF file that contains
both the code, your comments/conclusions and relevant output/figures, etc.\

In this project, we work with t he 2021 Australian daily weather database available online at       <http://www.bom.gov.au/climate/dwo/>. We will first read the data into R and then prepare data by going
through a number of issues so that a clean data set can be used for forecasting next day’s weather.
Finally, some simple exploratory data analysis (EDA) will be conducted.\

1.**(Data Input)** For simplicity, we shall focus on the weather data only at Canberra, which is
the Australian Capital Territory. Data for the latest month can be found at
<http://www.bom.gov.au/climate/dwo/IDCJDW2801.latest.shtml>
However, we will use the CSV text file. See the link by scrolling down the page. There are
many files like this at the same website for different months over the years.
With the following R code, we read in the 2021 daily weather data (available on Blackboard)
from January 1st till August 31st into one dataframe via a simple loop.\


```{r}
# READ DAILY WEATHER DATA IN 2021
setwd("C:/Users/chitr/OneDrive - University of Texas at El Paso/data_science/sem1-fall-2021/stat5474_datamining-dr_poko/hw/project-1")
dat <- NULL
i0 = NULL
current.month <- 9
for (i in 1:(current.month - 1)){
i0 <- ifelse(i<10, paste("0", i, sep=""), i) 
mth <- paste("2021", i0, sep="")
bom <- paste("IDCJDW2801.", mth, ".csv", sep="")
dat.i <- read.csv(bom, skip=6, check.names=FALSE,
na.strings = c("NA", "", " "), stringsAsFactors = FALSE)
dat.i[, 1] <-toupper(month.abb[i])
dat <- rbind(dat,dat.i)
}
```

Along the way, we have created one month variable with values {JAN, FEB, . .. AUG }. Report the
dimension of the data you obtain and include the first few data lines in your report.
``` {r}
dim(dat)
```

The data we created has 243 rows and 22 colums.

few data lines
``` {r}
head(dat)
```

2. **(Data Cleaning and Preparation)** We next clean the data and make it ready for the
subsequent analysis. Specifically, we shall address the following issues:
(a) First print out the (sorted) unique values or levels of every variable in the data set. You
may obtain the frequency table by treating all variables as if categorical. To do so, you may
consider using table(x, useNA="ifany"), where the useNA="ifany" option allows us to
see missing values.
``` {r}
#colnames(dat)[1] = "Month" # naming the first column
#var = colnames(dat)
#for (i in 1:length(var)) {table(dat[,i],useNA = "ifany")
lapply(dat,table)
#}
```

#``` {r}
#a = list()
#for (i in 1:length(names(dat))) {
# a[i]= table(dat[,i],useNA = "ifany")             }
#table(dat$`Maximum temperature (°C)`,useNA = "ifany")
#```

Then inspect for suspicious or problematic records. Comment on issues that you find out.
For example, you may find that variable called “9am wind speed (km/h)” has records
with value Calm. This value should be changed to 0 for the analysis purpose. What other
variables have similar problems?
(b) The variable called “Time of maximum wind gust” in the 10-th column may not be useful
for weather forecast purpose and its character-valued time info is difficult to enter any
model. So let’s remove it from the data set.
```{r} 
dat <- dat[, -c(10)]
```

(c) The variable names are too long. Rename the data set as follows.
``` {r} 
names(dat) <- c("Month", "Date", "MinTemp", "MaxTemp", "Rainfall",
"Evaporation", "Sunshine", "WindGustDir", "WindGustSpeed",
"Temp9am", "Humidity9am", "Cloud9am", "WindDir9am",
"WindSpeed9am", "Pressure9am", "Temp3pm", "Humidity3pm",
"Cloud3pm", "WindDir3pm", "WindSpeed3pm", "Pressure3pm")
dim(dat);
names(dat)
```
(d) Variables that have Calm as a value would be automatically treated by R as ‘character’ or
categorical, which is not the right way. For these variables, first change their Calm values
as 0 and then change their types into ‘numerical’ using the function as.numeric().
``` {r}
dat[dat == "Calm"]= as.numeric(0)
```

(e) Define a variable called RainToday based on Rainfall so that RainToday is 1 if Rainfall
is greater than 1mm and 0 otherwise. Namely, if it rains less than 1 mm in a day, then
we report that as no rain.
``` {r} 
dat$RainToday = ifelse(dat$Rainfall> 1,1,0)
```
Next, define a variable called RainTomorrow by shifting RainToday one day forward:
``` {r} 
dat$RainTomorrow <- c(dat$RainToday[2:nrow(dat)], NA)
```
Now the data is pretty much ready for some EDA and weather forest usage, where the
0-1 binary variable RainTomorrow is the target or response variable.

(f) Save a Rdata copy of the data set you have prepared using the save() function that you
obtain but do not submit it with your project report.
``` {r}
save(dat,file = "dat.Rdata")
```


3. **(Exploratory Data Analysis)** Perform some EDA of your choice on the data set that you
obtained. As a general principal, numerically you may consider two-way contingency table
plus χ2 test (or Fisher exact test) for independence to assess the association between a categorical variable with the binary outcome and consider two-sample t test (or the nonparametric
Wilcoxon rank sum test) to assess the association between a continuous variable with the binary
outcome. Graphical tools can be used as well, e.g., histogram and/or boxplot for continuous
2
variables and bar plot and/or mosaic plot for categorical variables (possibly grouped by the
binary outcome) among many other choices.

``` {r}
# chi-square test for selected catoregorical variables with binary output
chisq.test(dat$WindDir3pm,dat$RainTomorrow)
chisq.test(dat$WindSpeed9am,dat$RainTomorrow) # not significant
chisq.test(dat$Cloud9am,dat$RainTomorrow) # not significant
```

#``` {r}
# Fisherman test 
#fisher.test(dat$WindDir3pm,dat$RainTomorrow)
#fisher.test(dat$WindSpeed9am,dat$RainTomorrow)
#fisher.test(dat$Cloud9am,dat$RainTomorrow)

#```

``` {r}
# boxplot
boxplot( dat$RainTomorrow~dat$WindDir9am)
boxplot(dat$Rainfall~dat$WindDir3pm)
boxplot(dat$RainTomorrow~dat$Cloud9am)
boxplot(dat$RainTomorrow~dat$Month)
```
The deliverable consists of a l ist of t hree interesting findings that you have discovered. For example, one finding one can obtain by exploring t he association between month and RainTomorrow
is that the raining likelihood varies with months significantly (p-value = 0.05997 based on
Fisher's exact test and test size of 0.1) and April has the least rainy days (1 days with
RainTomorrow = 1).
``` {r}
tab <- table(dat$Month, dat$RainTomorrow, useNA="no"); tab
chisq.test(tab)
fisher.test(tab, simulate.p.value =TRUE)
```
Finally, remember to upload a single PDF file produced with R Markdown containing all code,
(relevant) output and plots, accompanying comments as well as summary of your findings and
conclusions.



